---
title: "A3: SPL Library Checkouts"
output: html_document
---

### The Decline of Libraries Due to COVID-19

The Coronavirus pandemic left the entire globe in confusion and fear, during its early stages in 2020. This forced much of the world to adapt, in hopes of instilling normalcy to our everyday lives. There were many institutions that were forced to undergo systematic changes. These fields include education, corporate entities, and many other institutions. We are able to note these changes, presently, in the ways with which we interact and use educational tools.

### Summary Information

```{r message=FALSE, warning=FALSE, code=readLines("summary.R"), include=FALSE}
```

The Seattle public Library has recently published a lot of its physical and digital book checkout data, dating back as far as 2005. In light of the recent pandemic, there are general trends among this data set that prove there has been a decline in physical book checkouts, while digital book checkouts have been steadily rising in frequency, and in monthly checkouts, since the year 2018.

Through the years of 2018 to 2022, there has been a steady decrease in the amount of physical books that have been checked out from the Seattle Public Library. In 2018, `num_physical_checkouts_2018` were checked out. In 2019, `num_physical_checkouts_2019` were checked out. In 2020, when the Coronavirus pandemic first began, `num_physical_checkouts_2020` were checked out. During the peak years of the pandemic, there were `num_physical_checkouts_2021` in 2021, and `num_physical_checkouts_2022`. Based on the data collected, there is a clear indication of a decrease in the number of books checked out, during the peak years of the Coronavirus pandemic.

The Seattle Public Library has also released the data regarding digital and ebook checkouts from 2018 to 2022. In 2018, there were `num_digital_checkouts_2018` digital books checked out. In 2019, there was a gradual increase in the number of digital books being checked out, bringing the number to `num_digital_checkouts_2019`. At the start of the pandemic, we noticed another gradual increase, bringing this figure to `num_digital_checkouts_2020`. During the peak years of the Coronavirus pandemic, there were `num_digital_checkouts_2021` and `num_digital_checkouts_2022` digital book checkouts. Based on this data, we can conclude that there has been a steady rise in the amount of digital books that Seattle residents have checked out, regardless of the global pandemic.



### The Dataset

The dataset that the Seattle Public Library published was last updated April 21, 2023. The dataset itself, was provided by the Seattle Public Library. The owner of the dataset is David Christensen. David Christensen is a Data Analyst for the Seattle Public Library, and has been for the past 12 years. He also works as avData Management Consultant.

The library data set that is being studied currently features data from 2017 to 2023. The data was collected by retrieving book checkout records from the Seattle Public library. This data was then cleaned and formatted to view in an organized fashion. Data, like the types found in this dataset are often used to study and observe what types of features and accommodations are continuously being used from a service, in order to make that service more accessible to the population, through innovation. In the data set, there are over 800,00 observations, with 12 distinct features. These features include the kinds of checkouts, physical or digital, the month and year of the checkout, the amount of checkouts for that date, the titles of the books checked out, the ISBN, of those books, and the publication year of those books. The dataset also features the format of the book that was checked out. This indicates if the book was an audio book, an ebook, a video dics, or a physical book.

An ethical question that arose while viewing and working with this data is: would it be ethical to retrieve data from hundreds of thousands of people, without their knowledge, for an educational study?



### First Trends Over Time Chart


```{r, echo=FALSE, warning=FALSE, message=FALSE, code=readLines("chart1.R")}
```
This chart shows the decrease in physical books being checked out over the course of various years.

### Second Trends Over Time Chart

```{r, echo=FALSE, message=FALSE, code=readLines("chart2.R")}
```
This chart shows the increase of digital books being checked out over the course of various years.

### Physical Book Checkouts 2017-2023


```{r, echo=FALSE, message=FALSE, code=readLines("chart3.R")}
```
This chart shows a faceted graph of both digital and physical book checkouts, over the course of various years. This chart demonstrates a side by side difference in how physical and digital books were checked out from 2017 to the beginning of 2023.
